---
title: "test"
author: "sn"
date: "2024-03-25"
output: pdf_document
---
```{r, echo=FALSE}
library(ggplot2)
library(tidyverse)
library(maps)
library(readr)
world <- map_data("world")
```

## Import data
```{r}
data <- read.csv("substance_use.csv")
head(data)
```


## In 2019, what region of the world has the highest rate of alcohol-related deaths among men aged 40-44?
```{r}
alcohol_death <- data[data$year == '2019' & data$cause == 'Alcohol use disorders' & data$sex == 'Male' & data$age == '40 to 44' & data$measure == 'Deaths', ]
head(alcohol_death)

highest_val <- alcohol_death[which.max(alcohol_death$val), ]
highest_region <- highest_val$location
print(highest_region)

```

## Looking at the prevalence of alcohol-related disease in the East Asia and Pacific region, how has this changed over time and in the different age groups? Is there a difference between men and women?

### select the East Asia and Pacific region
```{r}
Asia_Pacific <- data[data$location == 'East Asia & Pacific - WB' & data$cause == 'Alcohol use disorders' & data$measure == 'Prevalence', ]
```

### how the prevalence of alcohol-related disease change over time
```{r}
time <- ggplot(Asia_Pacific, aes(x = year, y = val, color = age, shape = sex)) + 
  geom_line() +
  geom_point()
print(time)
```

### differences between males and females
```{r}

Male <- Asia_Pacific[Asia_Pacific$sex == 'Male', ]
Female <- Asia_Pacific[Asia_Pacific$sex == 'Female', ]

shapiro.test(Male$val)
shapiro.test(Female$val)
wilcox.test(Male$val, Female$val, alternative = 'two.sided')
```

## Looking at the data from the United States, can you confirm an increase in the prevalence of diseases related to opioid use? What age group is the most affected?
```{r}
US <- data[data$location == 'North America' & data$cause == 'Opioid use disorders' & data$measure == 'Prevalence', ]

opioid1 <- ggplot(US[US$sex == 'Male', ], aes(x = year, y = val, color = age)) +
  geom_line() +
  geom_point() +
  ggtitle('Male')

opioid2 <- ggplot(US[US$sex == 'Female', ], aes(x = year, y = val, color = age)) +
  geom_line() +
  geom_point() +
  ggtitle('Female')

print(opioid1 + opioid2)




```

## Ask your own question

```{r}


map_data <- select(data, location, measure, cause, year, val, age, sex) 

alcohol_2019 <- map_data[map_data$measure == 'Prevalence' & map_data$cause == 'Alcohol use disorders' & map_data$year == '2019', ]


ggplot() +
  geom_map(data = world, map = world,
           aes(x = long, y = lat, map_id = region),
           fill = 'lightgrey', color = "black") +
 
  theme_void()



```



```{r}
South_Asia <- c('Afghanistan', 'Bangladesh', 'Bhutan', 'India', 'Maldives', 'Nepal', 'Pakistan', 'Sri Lanka')
North_America <- c('Bermuda', 'Canada', 'USA')

world <- world %>%
  mutate(location = ifelse(region %in% South_Asia, "South Asia - WB",
                     ifelse(region %in% North_America, "North America", "Other")))

merged_data <- merge(world, alcohol_2019, by = "location", all.x = TRUE)
mapdata <- select(merged_data, location, long, lat, measure, cause, year, sex, age, val)

target <- select(mapdata[mapdata$sex == 'Male' & mapdata$age == '30 to 34', ], location, long, lat, val)

ggplot() +
  geom_map(data = target, map = target,
           aes(x = long, y = lat, map_id = location),
           fill = "lightgray", color = "black") +
  geom_map(data = mapdata, map = mapdata,
           aes(fill = color, map_id = location),
           color = val) +
   scale_color_gradient(low = "lightblue", high = "darkblue") 
  theme_void()





```

```{r}
library(ggplot2)

# 添加一个 id 列到 target 数据框中
target$id <- 1:nrow(target)

# 绘制世界地图
p <- ggplot() +
  geom_map(data = world, map = world,
           aes(x = long, y = lat, map_id = region),
           fill = "lightgrey", color = "black") +
  theme_void()

# 在指定区域根据 val 值填充颜色
p +
  geom_map(data = target, map = world,
           aes(fill = val, map_id = id),
           color = "black") +
  scale_fill_gradient(low = "lightblue", high = "darkblue")


```

```{r}




```